\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-27 pldi/draft.tex
Teensy fixes

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
https://git.io/JfEgr
%% Author with single affiliation.
\author{
  Anonymous Author(s)\\
  \normalsize{Additional Anonymized Material: \href{https://git.io/JfwSq}{https://git.io/JfwSq}.}
}
%% \authornote{with author1 note}          %% \authornote is optional;
%%                                         %% can be repeated if necessary
%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%% \affiliation{
%%   \position{Position1}
%%   \department{Department1}              %% \department is recommended
%%   \institution{Institution1}            %% \institution is required
%%   \streetaddress{Street1 Address1}
%%   \city{City1}
%%   \state{State1}
%%   \postcode{Post-Code1}
%%   \country{Country1}                    %% \country is recommended
%% }
%% \email{first1.last1@inst1.edu}          %% \email is recommended

@@

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for 2--64$\times$-parallelism.}
      Different configurations per benchmark:
      (i) Par+Split: \ttt{eager} and general \ttt{split} enabled,
      (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (iii) Parallel: \ttt{eager} enabled (no \ttt{split}),
      (v) Blocking Eager, only blocking \ttt{eager} enabled (no \ttt{split}),
      (iv) No Eager: both \ttt{eager} and \ttt{split} disabled,
      (vi) Opt. Parallel, an optimized version of \ttt{bigrams}~\cf{ours}.
      Only relevant configurations are shown---\eg \ttt{grep} and \ttt{sort} do not benefit from \ttt{split}; \ttt{spell} and \ttt{bi-grams} do not see benefits without \ttt{split}.
      % TODO: WHY?
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
(Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the best \sys configuration, \ie \ttt{eager} enabled with \ttt{split} when it provides benefit, for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are 1.97, 3.5, 5.78, 8.83, 10.96, and 13.47, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}

@@

\heading{Discussion}
\sys achieves near-linear speedup on low-parallelism configurations.
This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).

No Eager and Blocking Eager depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}. These configurations perform worse than \sys with \ttt{eager} in all cases (except for a negligible difference in shortest-scripts) for the reasons described in \Cref{optimizer}. Adding \ttt{split} on top of that (Par + Split, Par + B. Split) further improves speedup for some scripts that have \tpur or \tnpu commands, and for the rest it does not affect performance. Finally, the optimized bi-grams script achieves significantly more benefits from \sys than its unoptimized counterpart, showing that familiarity with \sys's parallelizability classes goes a long way in terms of performance.

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}
\heading{Results}
\sys's parallelism is set to $16\times$, as many of these pipelines are relatively long (\S\ref{ours}, Discussion).
Their input was generated by multiplying their original input several times, enough to reach 10GB.
Fig.~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines.
Average speedup is $5.49\times$, median is $6.07\times$, and weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\pichskip{15pt}% Horizontal gap between picture and text
\parpic[r][t]{%
  \begin{minipage}{40mm}
    \includegraphics[width=\linewidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}%
    %% \captionof{figure}{This is a tiger.}
  \end{minipage}
}
\heading{Parallel Sort}
%% \kk{Explain what is the purpose of the
%%   microbenchmark. We want to establish that \sys scalability is as
%%   good as it could be.}
We first compare a single GNU \ttt{sort} optimized by \sys ($S_{PaSh}$) versus the same \ttt{sort} with the 
%% \begin{wrapfigure}[10]{r}{0.25\textwidth}
%%   \centering
%%   %% \begin{center}
%%     \includegraphics[width=0.25\textwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
%%   %% \end{center}
%% %   \vspace{-15pt}
%% % \caption{ \textbf{Comparison with \ttt{sort -{}-parallel}.} }
%% \end{wrapfigure}
\ttt{-{}-parallel} flag set ($S_{GNU}$). % , shown on the side.
% The results are shown on the side-figure below.
While \ttt{-{}-parallel} is not a general solution, the comparison serves to establish a baseline for \sys.
$S_{GNU}$'s parallelism is $2\times$ $S_{PaSh}$'s (\ie the rightmost plot point for $S_{GNU}$ is for \ttt{-{}-parallelism=128}), 
to account for \sys's additional merge processes.
% and as such we note that the measurements are harsh for \sys.

>>>>

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.
% https://git.io/JfEgr
%% Author with single affiliation.
\author{
  Anonymous Author(s)\\
  \normalsize{Additional Anonymized Material: \href{https://git.io/JfwSq}{https://git.io/JfwSq}.}
}
%% \authornote{with author1 note}          %% \authornote is optional;
%%                                         %% can be repeated if necessary
%% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
%% \affiliation{
%%   \position{Position1}
%%   \department{Department1}              %% \department is recommended
%%   \institution{Institution1}            %% \institution is required
%%   \streetaddress{Street1 Address1}
%%   \city{City1}
%%   \state{State1}
%%   \postcode{Post-Code1}
%%   \country{Country1}                    %% \country is recommended
%% }
%% \email{first1.last1@inst1.edu}          %% \email is recommended

@@

\heading{A Note on the Parallelism Factor}
\nv{
The discussion of the benchmarks often refers to the parallelism factor 
This is not necessarily the resulting parallelism achieved by \sys, which typically is above the level configured by the user.
This is because for each process (node in the graph), \sys creates multiple processes---aggregators, splitters, relays \etc
Some times the results give the impression of superlinear parallelism, but this has a simple explanation.
}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{\detokenize{./figs/tiling_throughput_scaleup.pdf}}
    \caption{
      \textbf{\sys's speedup for 2--6$\times$-parallelism.} 
      Different configurations per benchmark:
      (i) Par+Split: \ttt{eager} and general \ttt{split} enabled,
      (ii) Par+B.Split: \ttt{eager} and input-aware \ttt{split} enabled,
      (iii) Parallel: \ttt{eager} enabled (no \ttt{split}),
      (v) Blocking Eager, only blocking \ttt{eager} enabled (no \ttt{split}),
      (iv) No Eager: both \ttt{eager} and \ttt{split} disabled,
      (vi) Opt. Parallel, an optimized version of \ttt{bigrams}~\cf{ours}.
      Only relevant configurations are shown---\eg \ttt{grep} and \ttt{sort} do not benefit from \ttt{split}; \ttt{spell} and \ttt{bi-grams} do not see benefits without \ttt{split}.
      % TODO: WHY?
    }
    \vspace{-15pt}
    \label{fig:microbenchmarks}
\end{figure*}

@@

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
(Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are 1.97, 3.5, 5.78, 8.83, 10.96, and 13.47, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}

@@

\heading{Discussion}
\sys achieves near-linear speedup on low-parallelism configurations.
This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
An additional conclusion is that \sys's modular eager-relay significantly improves performance over the other alternatives (blocking eager or no eager, \S\ref{optimizer}).

@@

%% \TODO{An important point that needs to be made clear is that these are
%%   taken from the wild unmodified. And that Pash never leads to
%%   slowdown (except for the trivial head cases, where the slowdown is
%%   neligible).}
\heading{Results}
\sys's parallelism is set to $16\times$, as many of these pipelines are relatively long (\S\ref{ours}, Discussion).
Their input was generated by multiplying their original input several times, enough to reach 10GB.
Fig.~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for all 34 pipelines.
(\nv{Pipelines are referred by number, in both the plot and the discussion below.})
Average speedup is $5.49\times$, median is $6.07\times$, and weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\pichskip{15pt}% Horizontal gap between picture and text
\parpic[r][t]{%
  \begin{minipage}{40mm}
    \includegraphics[width=\linewidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}%
    %% \captionof{figure}{This is a tiger.}
  \end{minipage}
}
\heading{Parallel Sort}
%% \kk{Explain what is the purpose of the
%%   microbenchmark. We want to establish that \sys scalability is as
%%   good as it could be.}
We first compare a single GNU \ttt{sort} optimized by \sys ($S_{PaSh}$) versus the same \ttt{sort} with the 
%% \begin{wrapfigure}[10]{r}{0.25\textwidth}
%%   \centering
%%   %% \begin{center}
%%     \includegraphics[width=0.25\textwidth]{\detokenize{./figs/sort_baseline_comparison_scaleup.pdf}}
%%   %% \end{center}
%% %   \vspace{-15pt}
%% % \caption{ \textbf{Comparison with \ttt{sort -{}-parallel}.} }
%% \end{wrapfigure}
\ttt{-{}-parallel} flag set ($S_{GNU}$). % , shown on the side.
% The results are shown on the side-figure below.
While \ttt{-{}-parallel} is not a general solution, the comparison serves to establish a baseline for \sys.
$S_{GNU}$'s parallelism is configured to $2\times$ that of $S_{PaSh}$'s (\ie the rightmost plot point for $S_{GNU}$ is for \ttt{-{}-parallelism=127}), 
to account for \sys's additional merge processes.
% and as such we note that the measurements are harsh for \sys.
